---
title: "Plot_DNs_across_datasets"
author: "Tomke Stuerner"
date: "2024-09-11"
output: html_document
---
This script provides an example of how to retrieve annotations from our supplemental files and plot DNs from different datasets. 

Examples are the neuron images shown in Figure 1h.

Once only step to install necessary R packages. 
coconatfly gives access to main Drosophila connectomics datasets 
```{r, eval=FALSE}
# install natmanager if required
if (!requireNamespace("natmanager")) install.packages("natmanager")
natmanager::install('natverse')
natmanager::install(pkgs="coconatfly")
# set up token to access FANC data - will open your browser
fancr::fanc_set_token()
```

load R packages
```{r}
library(nat)
library(elmr)
library(fafbseg)
library(malevnc)
library(fancr)
library(dplyr)
```

Define which version should be used for each dataset. This is mostly here
so that people with access to the production versions of datasets still use
these default versions.
```{r}
FANC_version <- 851
FAFB_version <- 783
malevnc::choose_malevnc_dataset('MANC')
```
get the ids of the DN you would like to plot
Example DNa02 left side
```{r}
# define the type
type = "DNa02"
# read in SI files for data
FAFB_DNs=data.table::fread(file = '../Supplemental_files/Supplemental_file5_FAFB_DNs.tsv')
FANC_DNs =data.table::fread(file = '../Supplemental_files/Supplemental_file6_FANC_DNs.tsv')
MANC_DNs =data.table::fread(file = '../Supplemental_files/Supplemental_file7_MANC_DNs.tsv')

# pull ids for DNa02 from those supplemental files
DNa02_FAFB = FAFB_DNs |> 
  filter(type=="DNa02" & side == "L") |> 
  pull(root_id)

DNa02_FANC = FANC_DNs |> 
  filter(type=="DNa02" & side == "L") |> 
  pull(root_id)

DNa02_MANC = MANC_DNs |> 
  filter(type=="DNa02" & side == "L") |> 
  pull(bodyid)
```

plot DNa02 with the FAFB-Flywire brain surface mesh
```{r}
# load FAFB-Flywire mesh
DNa02_FAFB_mesh = read_cloudvolume_meshes(DNa02_FAFB)
# transform it into the template space
DNa02_FAFB_mesh_xf = xform_brain(DNa02_FAFB_mesh, sample = "FAFB14", reference = JFRC2013)
# transform the brain neuropil meshes into the same space as the DN
NP = xform_brain(JFRC2NP.surf, sample = "JFRC2", reference = JFRC2013)
# open a 3d scene
open3d(userMatrix = structure(c(1, 0, 0, 0, 0, -1, 0, 0, 0, 0, -1, 0, 0, 0, 0, 1), .Dim = c(4L,4L)), zoom = 0.556837737560272, windowRect = c(0L, 44L, 1686L, 1090L))
# chose the orientation, in this case an anterior view of the brain
nview3d("anterior")
# plot the neuron
plot3d(DNa02_FAFB_mesh_xf, col = "black")
# plot all neuropils in one shade of grey
plot3d(NP, alpha=0.04, add=TRUE, col = "grey")
# save a snapshot
dir.create("code/myplots")
rgl.snapshot(filename = "myplots/DNa02_L_brain.png")
# chose a different orientation, side view of the brain
nview3d("left")
```

plot with the brain neuropils that DNa02 primarily innervates: LAL and IPS
```{r}
NP = xform_brain(JFRC2NP.surf, sample = "JFRC2", reference = JFRC2013)
open3d(userMatrix = structure(c(1, 0, 0, 0, 0, -1, 0, 0, 0, 0, -1, 0, 0, 0, 0, 1), .Dim = c(4L,4L)), zoom = 0.556837737560272, windowRect = c(0L, 44L, 1686L, 1090L))
nview3d("anterior")
plot3d(DNa02_FAFB_mesh_xf, col = "black")
plot3d(NP, "LAL", alpha=0.4, col = "cyan")
plot3d(NP, "IPS", alpha=0.4, col = "lightcyan")
plot3d(NP, "SPS", alpha=0.4, col = "lightcyan")
plot3d(NP, alpha=0.04, add=TRUE, col = "grey")
```


plot DNa02 from MANC and FANC 
in the MANC space and with the MANC surface mesh
```{r}
# DNa02 mesh from MANC
DNa02_MANC_mesh = read_manc_meshes(DNa02_MANC)
mesh_px = DNa02_MANC_mesh/1e3
DNa02_MANC_mesh_sym = symmetric_manc(mesh_px)
names(DNa02_MANC_mesh_sym) = DNa02_MANC
# DNa02 mesh from FANC transformed into MANC space
DNa02_FANC_mesh  = read_fanc_meshes(DNa02_FANC)
FANC_mesh_m_xf=transform_fanc2manc(DNa02_FANC_mesh)
DNa02_FANC_mesh_sym = symmetric_manc(FANC_mesh_m_xf/1e3, mirror=F)
names(DNa02_FANC_mesh_sym) = DNa02_FANC
# MANC surface mesh of the VNC
MANCsym.surf=symmetric_manc(MANC.surf, mirror=F)

# plot
open3d(userMatrix = structure(c(1, 0, 0, 0, 0, 0, -1, 0, 0, 1, 0, 0, 0, 0, 0, 1), .Dim = c(4L,4L)), zoom = 0.613913536071777, windowRect = c(1792L, 45L, 2445L, 1079L))
# you can define an orientation using the function manc_view3d()
# manc_view3d('ventral')
plot3d(DNa02_MANC_mesh_sym,col = "red" ,lwd=2)
plot3d(DNa02_FANC_mesh_sym,col = "black" ,lwd=2)
plot3d(MANCsym.surf, alpha=0.04, add=TRUE, col = "grey")
```

plot with the VNC neuropils that DNa02 is primarily innervating: leg neuropils
```{r}
neuropil_colours <- c("LNp_T1_R"= "#D72000",
                        "LNp_T2_R"= "#F4B95A",
                        "LNp_T3_R"= "#EFEF46",
                        "LNp_T1_L"= "#D72000",
                        "LNp_T2_L"= "#F4B95A",
                        "LNp_T3_L"="#EFEF46")

reg_list_all= c("LNp_T1_R", "LNp_T2_R","LNp_T3_R","LNp_T1_L", "LNp_T2_L","LNp_T3_L")
open3d(userMatrix = structure(c(1, 0, 0, 0, 0, 0, -1, 0, 0, 1, 0, 0, 0, 0, 0, 1), .Dim = c(4L, 
4L)), zoom = 0.613913536071777, windowRect = c(1792L, 45L, 2445L, 1079L))
# manc_view3d('ventral')
plot3d(DNa02_MANC_mesh_sym,col = "red" ,lwd=2)
plot3d(DNa02_FANC_mesh_sym,col = "black" ,lwd=2)
for (i in 1:length(reg_list_all)) {
  x <- subset(MANCsym.surf, reg_list_all[i])
  plot3d(x,col = neuropil_colours[i], add=TRUE, alpha=0.2, lit = FALSE)
  assign(reg_list_all[i], x)}
plot3d(MANCsym.surf, alpha=0.04, add=TRUE, col = "grey")
```

plot with the VNC tract that DNa02 takes in the VNC: ITD
```{r}
# check which longitudinal tract has been assigned to the DN
FANC_DNs$long_tract[FANC_DNs$type==type]
# [1] "ITD" "ITD"
MANC_DNs$long_tract[MANC_DNs$type==type]
# [1] "ITD" "ITD"
tract_list_all= c("ITD_R", "ITD_L")
tract <- subset(MANC.tracts, tract_list_all)
tractsym = symmetric_manc(tract)

open3d(userMatrix = structure(c(1, 0, 0, 0, 0, 0, -1, 0, 0, 1, 0, 0, 0, 0, 0, 1), .Dim = c(4L, 
4L)), zoom = 0.613913536071777, windowRect = c(1792L, 45L, 2445L, 1079L))
# manc_view3d('ventral')
plot3d(DNa02_MANC_mesh_sym,col = "red" ,lwd=2)
# plot3d(DNa02_FANC_mesh_sym,col = "black" ,lwd=2)
plot3d(tractsym, col = "green")
plot3d(MANCsym.surf, alpha=0.04, add=TRUE, col = "grey")
```
