---
title: "FAFB_DNs_Fig3"
author: "Tomke Stuerner"
date: "2024-09-26"
output: html_document
---
This script shows examples of how to load and use the annotations provided in the Supplemental file5 FAFB_DNs sheet.
- plot input and output partners by class as in Fig. 3a
- plot the DN clusters by their primary brain neuropil as in Fig. 3c
- plot the clusters that are closest to gustatory sensory neurons as in Fig. 3e

```{r setup, include=FALSE}
library(fafbseg)
library(googlesheets4)
library(dplyr)
library(ggplot2)
library(tidyr)
library(forcats)
choose_segmentation("flywire31")
```

load the Supplemental file5 FAFB_DNs
```{r, include=FALSE}
FAFB_DNs = data.table::fread(file = '../Supplemental_files/Supplemental_file5_FAFB_DNs.tsv')
DN_ids = FAFB_DNs$root_id
```

upstream and downstream partners of DNs in FAFB-Flywire, weight>10
```{r, include=FALSE}
DN_US = flywire_partner_summary2(DN_ids, partners = "input", version = 783,threshold=10 )
DN_USsum <- aggregate(weight~super_class,DN_US,sum) 
DN_DS = flywire_partner_summary2(DN_ids, partners = "output", version = 783, threshold=10)
DN_DSsum <- aggregate(weight~super_class,DN_DS,sum) 
```

choose colors
```{r, include=FALSE}
delacroix = c("#C70E7B", "#FC6882", "#007BC3", "#54BCD1",
"#009F3F", "#8FDA04", "#AF6125", "#B25D91",
"#EF7C12", "#F4B95A", "#C23A4B", "#FBBB48", "#EFEF46", "#31D64D",
"#132157","#EE4244", "#D72000", "#1BB6AF", "#8B008B", "#551A8B","#808080")

names(delacroix) = c("magenta", "pink",
"blue", "cyan",
"darkgreen", "green",
"brown",
"mauve",
"darkorange", "orange",
"darkred", "darkyellow",
"yellow", "palegreen",
"navy","cerise",
"red", "marine",
"purple","darkpurple", "grey")

class_colours <- c("sensory" = delacroix[["green"]],
                   "sensory_ascending"  = delacroix[["darkgreen"]],
                   "descending" = delacroix[["pink"]],
                   "central" = delacroix[["cyan"]],
                   "ascending" = delacroix[["blue"]],
                   "motor" = delacroix[["purple"]],
                   "visual_projection" = delacroix[["brown"]],
                   "visual_centrifugal" = delacroix[["orange"]],
                   "optic" = delacroix[["red"]],
                   "endocrine" = "#A6A6A6")
```

plot input partners by class as in Fig. 3a
```{r, include=FALSE}
DN_USsum$col <- class_colours[match(DN_USsum$super_class,names(class_colours))]
DN_USsum$percent <- round(DN_USsum$weight*100/sum(DN_USsum$weight),1)

DN_USsum_pie<- ggplot(DN_USsum, aes(x="", y=percent, fill=super_class)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0) +
  scale_fill_manual(values=DN_USsum$col) +
  theme_minimal() +
  theme(axis.text.x=element_blank()) 

DN_USsum_pie
```
plot output partners by class as in Fig. 3a
```{r, include=FALSE}
DN_DSsum$col <- class_colours[match(DN_DSsum$super_class,names(class_colours))]
DN_DSsum$percent <- round(DN_DSsum$weight*100/sum(DN_DSsum$weight),1)

DN_DSsum_pie <- ggplot(DN_DSsum, aes(x="", y=percent, fill=super_class)) +
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) +
  scale_fill_manual(values=DN_DSsum$col) +  
  theme_minimal() +
  theme(axis.text.x=element_blank()) 

DN_DSsum_pie
```

plot the DN clusters by their primary brain neuropil
```{r, include=FALSE}
plot_neuropilgroup_sensorycluster = FAFB_DNs
plot_neuropilgroup_sensorycluster = subset(plot_neuropilgroup_sensorycluster, !plot_neuropilgroup_sensorycluster$cluster =="sensory descending")
plot_neuropilgroup_sensorycluster$number = 1
# how many DNs are in each cluster and each neuropilgroup
plot_neuropilgroup_sensorycluster2 = aggregate(plot_neuropilgroup_sensorycluster$number, 
                                               by = list(plot_neuropilgroup_sensorycluster$cluster, plot_neuropilgroup_sensorycluster$neuropilgroup), FUN = sum) 

names(plot_neuropilgroup_sensorycluster2) = c("cluster", "neuropilgroup", "number")
# normalise by the number of neurons in each cluster, shown in %
sum = aggregate(plot_neuropilgroup_sensorycluster$number, by = list(plot_neuropilgroup_sensorycluster$cluster), FUN = sum) 
names(sum) = c("cluster", "nr_of_neurons")
plot_neuropilgroup_sensorycluster2 = left_join(plot_neuropilgroup_sensorycluster2,sum)
plot_neuropilgroup_sensorycluster2$normtypes = (plot_neuropilgroup_sensorycluster2$number/plot_neuropilgroup_sensorycluster2$nr_of_neurons)*100
# plot heatmap
neuropilgroup_order = c(  "PRW_FLA","GNG","OC_LO","PS_LAL","multi","SMP_SLP",  "VES", "AVLP_AMMC_WED","PLP")

ggplot() + 
  geom_tile(data=plot_neuropilgroup_sensorycluster2, aes(x=cluster, y= factor(neuropilgroup, levels = neuropilgroup_order), fill=normtypes))+
  scale_fill_gradientn(colors=c("white","black","black")) +
  theme_minimal()+
  scale_x_discrete(name ="sensory clusters", 
                     limits=c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16"))+
  scale_y_discrete(name ="neuropilgroup")+
  theme(axis.text=element_text(size=13))
```


```{r, include=FALSE}
allsensory_ranking_type = data.table::fread(file = 'DN_ranking_df_783_sensory_combined_byDNtype.tsv')
# remove the sensory descending neurons
allsensory_ranking_type = subset(allsensory_ranking_type,!allsensory_ranking_type$cluster =="sensory descending")
gustatoryplot = subset(allsensory_ranking_type, select = c("type", "cluster", "gustatory"))
gustatoryplot %>%
  mutate(cluster = fct_reorder(cluster, gustatory, .fun='median')) %>%
  ggplot( aes(x=reorder(cluster, gustatory), y=gustatory, fill=cluster)) + 
  scale_fill_manual(values = c("1" = "#007BC3", "2" = "#007BC3", "3"="#8B008B", "4"="#8B008B", "5"="#132157" , "6" = "#54BCD1", "7"= "#8FDA04", "8"= "#8FDA04", "9"= "#8FDA04", "10"="#8B008B", "11"="#8B008B", "12"="#8B008B", 
                               "13"="#009F3F", "14"= "#8FDA04", "15"="#8B008B", "16"= "#8FDA04", "sensory descending" = "black"))+
  geom_boxplot()  +
  theme(legend.position="none") +
  theme(axis.title.x=element_blank(),axis.title.y=element_blank())+
  theme(axis.text=element_text(size=20))

```
