---
title: "DNa02"
author: "Tomke Stuerner"
date: "2024-09-10"
output: html_document
---

Read the following supplemental sheets by installing this branch of `coconatfly`.
```{r}
natmanager::install(pkgs = 'coconatfly@feature/more-fanc-meta', method = 'remotes')
```

Restart the session and load R packages
```{r}
library(coconatfly)
library(fafbseg)
library(readxl)
library(dplyr)
library(data.table)
library(fancr)
library(ggplot2)
library(plotly)
library(tidyr)
```

downstream connectivity of DNa02 in MANC and FANC
```{r}
# get all output partners of DNa02 in the MANC and FANC dataset
dna02ds=cf_partners(cf_ids('/DNa02', datasets = c("fanc", 'manc')), partners = 'out')
dna02ds$pre_id = as.character(dna02ds$pre_id)
# get all meta information for DNa02
DNa02_meta = cf_meta(cf_ids('/DNa02', datasets = c("fanc", 'manc')))
# add sides of the DN
DNa02_sides = subset(DNa02_meta, select = c(id, side))
names(DNa02_sides) = c("pre_id", "DN_side")
dna02ds = left_join(dna02ds,DNa02_sides)
# add side to dataset
dna02ds$dataset2 = paste0(dna02ds$dataset,"_", dna02ds$DN_side, sep= "")
dna02ds$npost = 1
```
calculate output percent of the DN and input percent to the receiving neuron
```{r}
# outputpercent
dna02ds_outputsum = aggregate(dna02ds$weight, by = list(dna02ds$pre_id), FUN = sum)
names(dna02ds_outputsum)= c("pre_id", "nrpresum")
dna02ds_outputsum$pre_id = as.character(dna02ds_outputsum$pre_id)
dna02ds = left_join(dna02ds, dna02ds_outputsum)
dna02ds$outputP = (dna02ds$weight/dna02ds$nrpresum)*100 

# input percent
dna02ds_inputsum = cf_partners(dna02ds$post_key, partners = 'in')
dna02ds_inputsum = aggregate(dna02ds_inputsum$weight, by = list(dna02ds_inputsum$post_id), FUN = sum)
names(dna02ds_inputsum)= c("post_id", "nrpostsum")
dna02ds_inputsum$post_id = as.character(dna02ds_inputsum$post_id)
dna02ds$post_id = as.character(dna02ds$post_id)
dna02ds = left_join(dna02ds, dna02ds_inputsum)
dna02ds$inputP = (dna02ds$weight/dna02ds$nrpostsum)*100 
```
add more meta data for neurons: for example soma_division (brain or neuromere)
```{r}
MANC_FANC_types = data.table::fread(file = '../Supplemental_files/Supplemental_file13_other_MANC_FANC_matching.tsv')
# update the FANC ids from the sheet to the newest version
MANC_FANC_types$root_id = fanc_latestid(MANC_FANC_types$root_id)
# add neuromere annotation for all neurons from our sheet, add missing types for MANC and FANC 
MANC_FANC_types = subset(MANC_FANC_types, select = c(manc_match_id,root_id, type, side, soma_division))
names(MANC_FANC_types)[3] = "missingtype"
MANC_FANC_types1 = subset(MANC_FANC_types, select = c(root_id, missingtype, soma_division))
names(MANC_FANC_types1)[1] = "post_id"
MANC_FANC_types2 = subset(MANC_FANC_types, select = c(manc_match_id, missingtype,soma_division))
names(MANC_FANC_types2)[1] = "post_id"
MANC_FANC_types = bind_rows(MANC_FANC_types1,MANC_FANC_types2)
dna02ds = left_join(dna02ds,MANC_FANC_types)
# add missing types
dna02ds$type[is.na(dna02ds$type) & !is.na(dna02ds$missingtype)] = dna02ds$missingtype[is.na(dna02ds$type) & !is.na(dna02ds$missingtype)] 
```
connection by side 
```{r}
dna02ds$type2 = paste0(dna02ds$type,"_", dna02ds$soma_division, sep = "")
# remove the ones we have a type
dna02ds = subset(dna02ds, !dna02ds$type2 == "NA_NA")
dna02dss = aggregate(list(dna02ds$weight, dna02ds$npost), by = list(dna02ds$type2, dna02ds$dataset2), FUN = mean)
names(dna02dss) = c("type2", "dataset2", "weight", "npost")

dna02dsw_side <- dna02dss |> 
  tidyr::pivot_wider(id_cols = c(type2), names_from = dataset2, 
                     values_from = c(weight,npost ), values_fill = 0)
```
Plot FANC R vs L
```{r}
fig <- dna02dsw_side %>% 
  filter(!is.na(type2)) %>% 
  plotly::plot_ly(x = ~ weight_fanc_L, y = ~ weight_fanc_R,
  type = "scatter", 
  mode = "markers",
  # Hover text:
  text = ~ paste0("type.post: ",type2, "\n"),
  marker = list(color = "black")) %>% 
  layout(title = 'Weight in FANC',
         xaxis = list(title = "Left"),
  yaxis = list(title = "Right")) %>% 
  add_lines(x = ~weight_fanc_L, y = ~weight_fanc_L, 
            line = list(color = 'grey', dash = "dash"), 
            showlegend = FALSE, 
            inherit = FALSE)
fig
```
Plot MANC R vs L
```{r}
fig2 <- dna02dsw_side %>% 
  filter(!is.na(type2)) %>% 
  plotly::plot_ly(x = ~ weight_manc_L, y = ~ weight_manc_R,
  type = "scatter", 
  mode = "markers",
  # Hover text:
  text = ~ paste0("type.post: ",type2, "\n"),
  marker = list(color = "black")) %>% 
  layout(title = 'Weight in MANC',
         xaxis = list(title = "Left"),
  yaxis = list(title = "Right")) %>% 
  add_lines(x = ~ weight_manc_L, y = ~ weight_manc_L, 
            line = list(color = 'grey', dash = "dash"), 
            showlegend = FALSE, 
            inherit = FALSE)
fig2
```
Plot FANC and MANC R vs L
```{r}
dna02dsw_long <- dna02dsw_side |> 
  filter(!is.na(type2)) |> 
  pivot_longer(cols = starts_with("weight_"), 
               names_to = c("dataset", "side"), 
               names_pattern = "weight_([^_]+)_([^_]+)",  # Use pattern to match dataset and side
               values_to = "weight") |> 
  pivot_wider(names_from = "side", values_from = "weight") |> 
  mutate(dataset = case_match(dataset,
             "fanc" ~ "FANC", 
             "manc" ~ "MANC"))

# Calculate Pearson correlation coefficient for each dataset
corr_fanc <- cor(dna02dsw_long$L[dna02dsw_long$dataset == "FANC"], 
                 dna02dsw_long$R[dna02dsw_long$dataset == "FANC"], use = "complete.obs")
corr_manc <- cor(dna02dsw_long$L[dna02dsw_long$dataset == "MANC"], 
                 dna02dsw_long$R[dna02dsw_long$dataset == "MANC"], use = "complete.obs")

fig3 <- dna02dsw_long %>%
  plot_ly(
    x = ~L, 
    y = ~R, 
    color = ~dataset, 
    colors = c("FANC" = "black", "MANC" = "red"),  # Colors for FANC and MANC
    text = ~paste0("type.post: ", type2, "\n"),  # Hover text
    type = "scatter", 
    mode = "markers"
  ) %>%
  
  # Add axis labels
  layout(
    xaxis = list(title = "Left Side Weight"),
    yaxis = list(title = "Right Side Weight"),
    legend = list(title = list(text = "Dataset")),
    showlegend = TRUE,
    annotations = list(
      # Add Pearson correlation annotation for FANC
      list(
        x = 1, y = 1, text = paste0("FANC Cor: ", round(corr_fanc, 2)),
        xref = "paper", yref = "paper", showarrow = FALSE,
        xanchor = 'right', yanchor = 'top', font = list(size = 12, color = "black")
      ),
      # Add Pearson correlation annotation for MANC
      list(
        x = 1, y = 0.95, text = paste0("MANC Cor: ", round(corr_manc, 2)),
        xref = "paper", yref = "paper", showarrow = FALSE,
        xanchor = 'right', yanchor = 'top', font = list(size = 12, color = "red")
      )
    )
  )

# Add a reference line for slope = 1
fig3 <- fig3 %>%
  add_lines(
    x = ~ L,
    y = ~ L,
    line = list(dash = "dash", color = "gray"),
    inherit = FALSE,  # Ensures the line is handled separately
    showlegend = FALSE
  )
fig3
```
To compare the top partners across the two datasets:
- we took types that were above the 1% threshold of DN total output

Plot FANC vs. MANC in mean weight
```{r}
# only take types of neurons that receive more than 1% of DNa02 output  
dna02_output_types = unique(dna02ds$type[dna02ds$outputP>1])
dna02ds2 = subset(dna02ds, dna02ds$type %in% dna02_output_types)
dna02ds2 = subset(dna02ds2, !is.na(dna02ds2$soma_division))
dna02dss2 = aggregate(list(dna02ds2$inputP, dna02ds2$weight, dna02ds2$npost), by = list(dna02ds2$type2, dna02ds2$dataset), FUN = mean)
names(dna02dss2) = c("type2", "dataset", "weightP", "weight", "npost")

dna02dsw_dataset <- dna02dss2 |> 
  tidyr::pivot_wider(
    id_cols = c(type2), 
    names_from = dataset, 
    values_from = c(weightP, weight, npost), 
    values_fill = 0
  )

# Create the scatter plot with Plotly
fig4 <- dna02dsw_dataset |> 
  filter(!is.na(type2)) |> 
  mutate(
    label = paste0(
      "type.post: ", type2, "\n", 
      "npost_fanc: ", npost_fanc, "\n",  # Adjust based on column names after pivot
      "npost_manc: ", npost_manc  # Adjust based on column names after pivot
    )
  ) %>%
  plot_ly(
    x = ~weight_manc,  # Use the MANC weight column
    y = ~weight_fanc,  # Use the FANC weight column
    text = ~label,  # Hover text
    type = "scatter",
    mode = "markers",
    marker = list(color = 'black')  # Customize marker color
  ) %>%
  
  # Add diagonal line (y = x) with slope = 1
  add_lines(
    x = ~weight_manc,
    y = ~weight_manc,
    line = list(dash = "dash", color = "gray"),
    inherit = FALSE,  # Ensures the line is handled separately
    showlegend = FALSE
  ) %>%
  # Layout for the axes and plot style
  layout(
    xaxis = list(title = "MANC Weight"),
    yaxis = list(title = "FANC Weight"),
    title = "Scatter Plot: MANC vs FANC Weights",
    paper_bgcolor = 'white',  # Background color outside the plot
    plot_bgcolor = 'white',   # Background color inside the plot
    font = list(family = "Arial", size = 12, color = 'black')  # Customize font
  )
fig4
```

Plot FANC vs. MANC in percent input to the receiving neurons
- to remove issues due to reconstruction state of the neurons
```{r}
fig5 <- dna02dsw_dataset |> 
  filter(!is.na(type2)) |> 
  mutate(
    label = paste0(
      "type.post: ", type2, "\n", 
      "npost_fanc: ", npost_fanc, "\n",  # Adjust based on actual column names after pivot
      "npost_manc: ", npost_manc  # Adjust based on actual column names after pivot
    )
  ) %>%
  plot_ly(
    x = ~weightP_manc,  # Adjust based on the column names for MANC weightP
    y = ~weightP_fanc,  # Adjust based on the column names for FANC weightP
    text = ~label,  # Hover text
    type = "scatter",
    mode = "markers",
    marker = list(color = 'black')  # Customize marker color
  ) %>%
  
  # Add diagonal line (slope = 1)
  add_lines(
    x = ~weightP_manc,
    y = ~weightP_manc,
    line = list(dash = "dash", color = "gray"),
    inherit = FALSE,  # Ensures the line is handled separately
    showlegend = FALSE
  ) %>%
  
  # Layout for the axes and plot style
  layout(
    xaxis = list(title = "MANC WeightP"),
    yaxis = list(title = "FANC WeightP"),
    title = "Scatter Plot: MANC vs FANC WeightP",
    paper_bgcolor = 'white',  # Background color outside the plot
    plot_bgcolor = 'white',   # Background color inside the plot
    font = list(family = "Arial", size = 12, color = 'black')  # Customize font
  )
fig5
```
For Figure 5 we plot the connectivity for the DNa02 top 4 output types: IN08A006, IN19A003, IN13B001, IN07B006, w-cHIN_T2
