---
title: "DNa02"
author: "Tomke Stuerner"
date: "2024-09-10"
output: html_document
---

Read the following supplemental sheets by installing this branch of `coconatfly`.
```{r}
natmanager::install(pkgs = 'coconatfly@feature/more-fanc-meta')
```

Restart the session and load R packages
```{r}
library(coconatfly)
library(fafbseg)
library(readxl)
library(dplyr)
library(data.table)
library(fancr)
library(ggplot2)
```

downstream connectivity of DNa02 in MANC and FANC
```{r}
# get all output partners of DNa02 in the MANC and FANC dataset
dna02ds=cf_partners(cf_ids('/DNa02', datasets = c("fanc", 'manc')), partners = 'out')
dna02ds$pre_id = as.character(dna02ds$pre_id)
# get all meta information for DNa02
DNa02_meta = cf_meta(cf_ids('/DNa02', datasets = c("fanc", 'manc')))
# add sides of the DN
DNa02_sides = subset(DNa02_meta, select = c(id, side))
names(DNa02_sides) = c("pre_id", "DN_side")
dna02ds = left_join(dna02ds,DNa02_sides)
# add side to dataset
dna02ds$dataset2 = paste0(dna02ds$dataset,"_", dna02ds$DN_side, sep= "")
dna02ds$npost = 1
```
calculate output percent of the DN and input percent to the receiving neuron
```{r}
# outputpercent
dna02ds_outputsum = aggregate(dna02ds$weight, by = list(dna02ds$pre_id), FUN = sum)
names(dna02ds_outputsum)= c("pre_id", "nrpresum")
dna02ds_outputsum$pre_id = as.character(dna02ds_outputsum$pre_id)
dna02ds = left_join(dna02ds, dna02ds_outputsum)
dna02ds$outputP = (dna02ds$weight/dna02ds$nrpresum)*100 

# input percent
dna02ds_inputsum = cf_partners(dna02ds$post_key, partners = 'in')
dna02ds_inputsum = aggregate(dna02ds_inputsum$weight, by = list(dna02ds_inputsum$post_id), FUN = sum)
names(dna02ds_inputsum)= c("post_id", "nrpostsum")
dna02ds_inputsum$post_id = as.character(dna02ds_inputsum$post_id)
dna02ds = left_join(dna02ds, dna02ds_inputsum)
dna02ds$inputP = (dna02ds$weight/dna02ds$nrpostsum)*100 
```
add more meta data for neurons: for example soma_devision (brain or neuromere)
```{r}
MANC_FANC_types = data.table::fread(file = '../Supplemental_files/Supplemental_file13_other_MANC_FANC_matching.tsv')
MANC_FANC_types$root_id = fanc_latestid(MANC_FANC_types$root_id)
MANC_FANC_types = subset(MANC_FANC_types, select = c(manc_match_id,root_id, type, side, soma_division))
# add neuromere annotation for all neurons from our sheet, add missing types
MANC_FANC_types1 = subset(MANC_FANC_types, select = c(root_id, missingtype, soma_division))

### TO CORRECT
# Error in eval(substitute(select), nl, parent.frame()) : 
#   object 'missingtype' not found

names(MANC_FANC_types1)[1] = "post_id"
MANC_FANC_types2 = subset(MANC_FANC_types, select = c(manc_match_id, missingtype,soma_division))
names(MANC_FANC_types2)[1] = "post_id"
MANC_FANC_types = bind_rows(MANC_FANC_types1,MANC_FANC_types2)
dna02ds = left_join(dna02ds,MANC_FANC_types)
# add missing types
dna02ds$type[is.na(dna02ds$type) & !is.na(dna02ds$missingtype)] = dna02ds$missingtype[is.na(dna02ds$type) & !is.na(dna02ds$missingtype)] 
```
connection by side 
```{r}
dna02ds$type2 = paste0(dna02ds$type,"_", dna02ds$soma_division, sep = "")
# remove the ones we have a type
dna02ds = subset(dna02ds, !dna02ds$type2 == "NA_NA")
dna02dss = aggregate(list(dna02ds$weightP, dna02ds$weight, dna02ds$npost), by = list(dna02ds$type2, dna02ds$dataset2), FUN = mean)
names(dna02dss) = c("type2", "dataset2", "weightP", "weight", "npost")

dna02dsw_side <- dna02dss |> 
  tidyr::pivot_wider(id_cols = c(type2), names_from = dataset2, 
                     values_from = c(weightP,weight,npost ), values_fill = 0)
```
Plot FANC R vs L
```{r}
dna02dsw_side |> 
  filter(!is.na(type2)) |> 
  mutate(label=paste0("type.post: ",type2, "\n")) |> 
  ggplot(aes(x=weight_fanc_L, y=weight_fanc_R, label=label)) + 
  geom_point()+
  geom_abline(slope = 1)->p
plotly::ggplotly(p)
```
Plot MANC R vs L
```{r}
dna02dsw_side |> 
  filter(!is.na(type2)) |> 
  mutate(label=paste0("type.post: ",type2, "\n")) |> 
  ggplot(aes(x=weight_manc_L, y=weight_manc_R, label=label)) + 
  geom_point()+
  geom_abline(slope = 1)->p
plotly::ggplotly(p)
```
Plot FANC and MANC R vs L
```{r}
dna02dsw_long <- dna02dsw_side |> 
  filter(!is.na(type2)) |> 
  pivot_longer(cols = starts_with("weight_"), 
               names_to = c("dataset", "side"), 
               names_pattern = "weight_([^_]+)_([^_]+)",  # Use pattern to match dataset and side
               values_to = "weight") |> 
  pivot_wider(names_from = "side", values_from = "weight") |> 
  mutate(dataset = ifelse(dataset == "fanc", "FANC", "MANC"))

# Calculate Pearson correlation coefficient for each dataset
corr_fanc <- cor(dna02dsw_long$L[dna02dsw_long$dataset == "FANC"], 
                 dna02dsw_long$R[dna02dsw_long$dataset == "FANC"], use = "complete.obs")
corr_manc <- cor(dna02dsw_long$L[dna02dsw_long$dataset == "MANC"], 
                 dna02dsw_long$R[dna02dsw_long$dataset == "MANC"], use = "complete.obs")

p_combined <- dna02dsw_long |> 
  ggplot(aes(x = L, y = R, color = dataset, label = paste0("type.post: ", type2, "\n"))) + 
  geom_point() +
  geom_abline(slope = 1, linetype = "dashed") +  # Reference line for slope=1
  labs(x = "Left Side Weight", y = "Right Side Weight", color = "Dataset") +
  scale_color_manual(values = c("FANC" = "black", "MANC" = "red")) +  # Colors for FANC and MANC
  theme_minimal() +
  
  # Add Pearson correlation coefficient annotations
  annotate("text", x = Inf, y = Inf, label = paste0("FANC Cor: ", round(corr_fanc, 2)), 
           color = "black", hjust = 2, vjust = 2, size = 5) +
  annotate("text", x = Inf, y = Inf, label = paste0("MANC Cor: ", round(corr_manc, 2)), 
           color = "red", hjust = 2, vjust = 3, size = 5)

# Convert to plotly for interactivity
plotly::ggplotly(p_combined)
```
To compare the top partners across the two datasets:
- we took types that were above the 1% threshold of DN total output

Plot FANC vs. MANC in mean weight
```{r}
# only take types of neurons that receive more than 1% of DNa02 output  
dna02_output_types = unique(dna02ds$type[dna02ds$outputP>1])
dna02ds2 = subset(dna02ds, dna02ds$type %in% dna02_output_types)
dna02ds2 = subset(dna02ds2, !is.na(dna02ds2$soma_division))
dna02dss2 = aggregate(list(dna02ds2$inputP, dna02ds2$weight, dna02ds2$npost), by = list(dna02ds2$type2, dna02ds2$dataset), FUN = mean)
names(dna02dss2) = c("type2", "dataset", "weightP", "weight", "npost")

dna02dsw_dataset <- dna02dss2 |> 
  tidyr::pivot_wider(id_cols = c(type2), names_from = dataset, 
                     values_from = c(weightP,weight,npost ), values_fill = 0)
dna02dsw_dataset |> 
  filter(!is.na(type2)) |> 
  mutate(label=paste0("type.post: ",type2, "\n", 
                     "npost_fanc:", npost_fanc, "\n",
                     "npost_manc:", npost_manc)) |> 
  ggplot(aes(x=weight_manc, y=weight_fanc, label=label)) + 
  geom_point()+
  theme_minimal() +
  geom_abline(slope = 1)->p
plotly::ggplotly(p)
```
Plot FANC vs. MANC in percent input to the receiving neurons
- to remove issues due to reconstruction state of the neurons
- for Figure 5 we plot the connectivity for the DNa02 top 4 output types: IN08A006, IN19A003, IN13B001, IN07B006, w-cHIN_T2

```{r}
dna02dsw_dataset |> 
  filter(!is.na(type2)) |> 
  mutate(label=paste0("type.post: ",type2, "\n", 
                     "npost_fanc:", npost_fanc, "\n",
                     "npost_manc:", npost_manc)) |> 
  ggplot(aes(x=weightP_manc, y=weightP_fanc, label=label)) + 
  geom_point()+
  theme_minimal() +
  geom_abline(slope = 1)->q
plotly::ggplotly(q)
```
