---
title: "Dimorphic_neuron_partners_manc"
author: "Tomke Stuerner"
date: "2024-09-26"
output: html_document
---

This script compares the number of partners of sexually dimorphic or male-specific DNs and ANs with partners of non-dimorphic DNs and ANs.

For this it pulls the annotations from the Supplemental files and used the malevnc package to query input and output partners.


```{r setup, include=FALSE}
library(malevnc)
library(data.table)
library(dplyr)
```

load the required supplemental files
```{r , include=FALSE}
dimorphic_DNs = data.table::fread(file = '/Users/tomke/Documents/dev/2023neckconnective/Supplemental_files/Supplemental_file14_dimorphic_DNs.tsv')
all_DNs <- data.table::fread(file = '/Users/tomke/Documents/dev/2023neckconnective/Supplemental_files/Supplemental_file7_MANC_DNs.tsv')
dimorphic_ANs<- data.table::fread(file = '/Users/tomke/Documents/dev/2023neckconnective/Supplemental_files/Supplemental_file15_dimorphic_ANs.tsv')
all_ANs <- data.table::fread(file = '/Users/tomke/Documents/dev/2023neckconnective/Supplemental_files/Supplemental_file11_MANC_ANs.tsv')
```
count how many DN and AN types are sexually dimorphic and male-specific
define 
```{r , include=FALSE}
# there are 10 male-specific DNs, 30 sexually dimorphic DNs and 1285 non-dimorphic in the MANC dataset
DNspec_m = subset(dimorphic_DNs, dimorphic_DNs$sex_specific_dimorphic %in% c("male specific"))
DNdim_both = subset(dimorphic_DNs, dimorphic_DNs$sex_specific_dimorphic == "dimorphic morphology")
non_dim_dn_manc = subset(all_DNs, !all_DNs$bodyid %in% dimorphic_DNs$MANC_bodyid)
# there are 76 male-specific ANs, 59 sexually dimorphic ANs and 1865 non-dimorphic in the MANC dataset
ANspec_m = subset(dimorphic_ANs, dimorphic_ANs$sex_specific_dimorphic  == "male specific")
ANdim_both = subset(dimorphic_ANs, dimorphic_ANs$sex_specific_dimorphic  == "dimorphic morphology")
non_dim_an_manc = subset(all_ANs, !all_ANs$bodyid %in% dimorphic_ANs$bodyid)
```

DN output partner numbers in MANC, thesholded at a weight >10
```{r , include=FALSE}
#partners of sexually dimorphic and male-specific DNs
DNdim_spec_manc <- c(DNspec_m$MANC_bodyid, DNdim_both$MANC_bodyid)
DNdim_spec_manc.ds <- manc_connection_table(DNdim_spec_manc, partners = 'out', moredetails=T)
DNdim_spec_manc.ds.sel <- aggregate(weight~partner,DNdim_spec_manc.ds,sum) 
length(unique(DNdim_spec_manc.ds.sel$partner[DNdim_spec_manc.ds.sel$weight>=10]))
# 2284
#partners of sexually dimorphic DNs
DNdim_manc.ds <- manc_connection_table(DNdim_both$MANC_bodyid, partners = 'out', moredetails=T)
DNdim_manc.ds.sel <- aggregate(weight~partner,DNdim_manc.ds,sum) 
length(unique(DNdim_manc.ds.sel$partner[DNdim_manc.ds.sel$weight>=10]))
# 1812
#partners of male-specific DNs
DNspec_manc.ds <- manc_connection_table(DNspec_m$MANC_bodyid, partners = 'out', moredetails=T)
DNspec_manc.ds.sel <- aggregate(weight~partner,DNspec_manc.ds,sum) 
length(unique(DNspec_manc.ds.sel$partner[DNspec_manc.ds.sel$weight>=10]))
# 637
#partners of non sexually dimorphic or male-specific DNs
non_dim_spec_manc.ds <- manc_connection_table(non_dim_dn_manc$bodyid, partners = 'out', moredetails=T)
non_dim_spec_manc.ds.sel <- aggregate(weight~partner,non_dim_spec_manc.ds,sum) 
length(unique(non_dim_spec_manc.ds.sel$partner[non_dim_spec_manc.ds.sel$weight>=10]))
# 18116
```

AN input partner numbers in MANC, thesholded at a weight >10
```{r , include=FALSE}
#partners of sexually dimorphic and male-specific ANs
ANdim_spec_manc <- c(ANspec_m$bodyid, ANdim_both$bodyid)
ANdim_spec_manc.us <- manc_connection_table(ANdim_spec_manc, partners = 'in', moredetails=T)
ANdim_spec_manc.us.sel <- aggregate(weight~partner,ANdim_spec_manc.us,sum)
length(unique(ANdim_spec_manc.us.sel$partner[ANdim_spec_manc.us.sel$weight>=10]))
# 3476
#partners of sexually dimorphic ANs
ANdim_manc.us <- manc_connection_table(ANdim_both$bodyid, partners = 'in', moredetails=T)
ANdim_manc.us.sel <- aggregate(weight~partner,ANdim_manc.us,sum) 
length(unique(ANdim_manc.us.sel$partner[ANdim_manc.us.sel$weight>=10]))
#1913
#partners of male-specific ANs
ANspec_manc.us <- manc_connection_table(ANspec_m$bodyid, partners = 'in', moredetails=T)
ANspec_manc.us.sel <- aggregate(weight~partner,ANspec_manc.us,sum) 
length(unique(ANspec_manc.us.sel$partner[ANspec_manc.us.sel$weight>=10]))
# 2204
#partners of non sexually dimorphic or male-specific ANs
non_dim_spec_an_manc.us <- manc_connection_table(non_dim_an_manc$bodyid, partners = 'in', moredetails=T)
non_dim_spec_an_manc.us.sel <- aggregate(weight~partner,non_dim_spec_an_manc.us,sum) 
length(unique(non_dim_spec_an_manc.us.sel$partner[non_dim_spec_an_manc.us.sel$weight>=10]))
# 25056
```
How many partner do DNdim_spec_manc and ANdim_spec_manc share? 
```{r , include=FALSE}
DNdim_spec_manc.ds.sel10 = unique(DNdim_spec_manc.ds.sel$partner[DNdim_spec_manc.ds.sel$weight>=10])
ANdim_spec_manc.us.sel10 = unique(ANdim_spec_manc.us.sel$partner[ANdim_spec_manc.us.sel$weight>=10])
common_partners_dimspec = intersect(DNdim_spec_manc.ds.sel10, ANdim_spec_manc.us.sel10) 
length(common_partners_dimspec)
# share 1007 partners, represents 44% of the DN output partners and 29% of the AN input partners
```


