---
title: "R Notebook"
output: html_notebook
---

This script may be interesting for external users but is primarily intended for
internal use in preparing annotations in a more convenient form for end
users.

To get started you need to install some R packages. 

```{r, eval=FALSE}
if(!requireNamespace('natmanager'))
  install.packages('natmanager')
natmanager::install(pkgs = 'coconatfly')
natmanager::install(pkgs = 'fafbseg')
# one more package ...
install.packages('readxl')
```

We recommend installing `coconatfly` since that brings everything else you'll
need with it. If you ever need to upgrade, you can just run `natmanager::install` again. fafbseg must be up to date so we'll also install that directly.

To access the annotations we'll use a function in the fafbseg package (which 
we'll acess with `fafbseg::`) and we'll likely do additional manipulation with coconatfly.

```{r}
library(coconatfly)
```

You can read the annotation files:
```{r}
readxl::excel_sheets(fafbseg::flywire_sirepo_file_memo('https://github.com/flyconnectome/2023neckconnective/supplemental_files/Supplemental_file_1_Seed_planes.xlsx'))
```

```{r}
read_all_sheets <- function(f) {
  sheets=readxl::excel_sheets(f)
  sapply(sheets, function(s) readxl::read_excel(f, sheet=s, guess_max=10e3), simplify = F)
}
```

```{r}
si1=read_all_sheets(fafbseg::flywire_sirepo_file_memo('https://github.com/flyconnectome/2023neckconnective/supplemental_files/Supplemental_file_1_Seed_planes.xlsx'))
si1
```

```{r}
si2=read_all_sheets(fafbseg::flywire_sirepo_file_memo('https://github.com/flyconnectome/2023neckconnective/supplemental_files/Supplemental_file_2_Typing_and_matching.xlsx'))
si2
```

si2 is the more interesting one. Let's take a look:

```{r}
names(si2)
```

```{r}
si2$DN_identification
```

The identification sheets provide details/evidence for light level identities.

```{r}
si2$DN_identification
```

## Uniform FANC annotations
```{r}
si2$FANC_DNs
```

Now let's say we're looking at FANC neurons, we might want to collect up the ids

```{r}
fanc_sheets=grep("FANC", names(si2), value = T)
```

```{r}
sapply(si2[fanc_sheets], colnames)
```
There are lots of useful data here but it will be hard to use for general type
information e.g. for coconatfly as we need uniform fields names across
the four sheets.

We'll need to do some processing

```{r}
library(dplyr)
```


First define a function to generate nuemric groups for neurons inferred to be
of the same type.

```{r}
type2group <- function(df, badtypes=character()) {
  ids=fafbseg::flywire_ids(df$root_id, integer64 = T)
  df2 <- df %>%
    filter(!duplicated(.data$root_id)) %>%
    filter(!.data$type %in% badtypes) %>%
    group_by(.data$type) %>%
    mutate(group = as.character(suppressWarnings(min(fafbseg::flywire_ids(.data$root_id, integer64 = T))))) %>%
    ungroup()
  res = fafbseg::flywire_ids(df2$group[match(ids, df2$root_id)], integer64 = F)
}
```


Now let's process each of the four sheets in turn with some specialised logic
to produce uniform results.

```{r}
fanc_sheets2=list()
fanc_sheets2[['FANC_DNs']] <- si2$FANC_DNs %>%
  mutate(lineage=NA_character_, soma_division="brain") %>%
  mutate(manc_group=fafbseg::flywire_ids(manc_group, integer64 = F)) %>%
  mutate(group=fafbseg::flywire_ids(match_group, integer64 = F)) %>%
  select(root_id, supervoxel_id, side, soma_division, class, type, lineage, group, manc_group)

fanc_sheets2[['FANC_ANs']] <- si2$FANC_ANs %>%
  rename(lineage=hemilineage, root_id=rootID, supervoxel_id=svID, soma_division=soma_neuromere) %>%
  mutate(manc_group=fafbseg::flywire_ids(manc_group, integer64 = F)) %>%
  mutate(group=type2group(.)) %>%
  select(root_id, supervoxel_id, side, soma_division, class, type, lineage, group, manc_group)

fanc_sheets2[['FANC_SAs']] <- si2$FANC_SAs %>%
  rename(supervoxel_id=svID) %>%
  mutate(manc_group=fafbseg::flywire_ids(NA, integer64 = F)) %>%
  mutate(group=type2group(.)) %>%
  mutate(lineage=NA_character_) %>%
  mutate(soma_division='body') %>%
  select(root_id, supervoxel_id, side, soma_division, class, type, lineage, group, manc_group)

fanc_sheets2[['other_MANC_FANC_matching']] <- si2$other_MANC_FANC_matching %>%
  rename(root_id=rootID, soma_division=soma_neuromere, side=soma_side) %>%
  # mutate(root_id=fanc_latestid(root_id)) %>%
  mutate(manc_group=fafbseg::flywire_ids(NA, integer64 = F)) %>%
  mutate(group=type2group(.)) %>%
  mutate(supervoxel_id=NA_character_, lineage=NA_character_) %>%
  select(root_id, supervoxel_id, side, soma_division, class, type, lineage, group, manc_group)

fanc_sheets2
```

Ok now let's bind all these sheets together and update the ids to the most
recent materialisation version which is what coconatfly uses for connectivity.
```{r}
fanc_anns <- dplyr::bind_rows(fanc_sheets2)
fanc_anns <- fanc_anns %>%
  mutate(
    root_id = fancr::with_fanc(
      fafbseg::flywire_updateids(root_id, supervoxel_id, version = 'latest'))
    ) %>%
  mutate(root_id=fancr::fanc_latestid(root_id, version = 'latest'))
```
Finally we might like to find the cell 

```{r}
fanc_anns |> View()
```

```{r}
fanc_anns2 <- fanc_anns %>%
  mutate(cell_id=fancr::fanc_cellid_from_segid(root_id, version = 'latest'))
fanc_anns2 |> 
  add_count(cell_id) |> 
  mutate(n=case_when(is.na(cell_id) ~ NA, T ~ n)) |> 
  count(n)
```

```{r}
fanc_anns2 |> 
  count(soma_division)
```

```{r}
fanc_anns2
```

```{r}
fanc_anns2 |> 
  add_count(root_id) |> 
  filter(n>1) |> 
  knitr::kable()
```

```{r}
fanc_anns2 |> 
  add_count(root_id) |> 
  filter(n>1 & grepl("sensory_ascending", class)) |> 
  select(root_id, supervoxel_id, class, type) |> 
  knitr::kable()

```

```{r}
fanc_anns2 |> 
  add_count(root_id) |> 
  filter(n>1 & !grepl("sensory_ascending", class)) |> 
  select(root_id, class, type, side, cell_id) |> 
  knitr::kable()

```

